### 2014-04-09 paper/designspace.tex
bjoern's edits through end of tools section

Open pipes may be used to create electronic circuitry.  For example, an open pipe filled with conductive paint and connected to a battery can provide power to an LED (see Figure \ref{fig:openings}A).
%
Semi-enclosed pipes are capped on one side. Caps can for example be used to create tactile output. By fabricating a cap from a malleable material and attaching a pump to the open ended, haptic feedback can be generated by raising and depressing the flexible cap (see Figure \ref{fig:openings}C).
%
A fully enclosed pipe or cavities are entirely internal to the object.  Fully enclosed pipes may be used as weight modifiers and for object identification, as in Figure \ref{fig:openings}D.  They can also contain for water or particles that would otherwise fall out.\bjoern{how would you get the particles inside?}

@@

\subsection{Pipe Topologies}
Pipe network topologies enable splitting or mixing of media (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out configurations; we use star topologies to create different conductive paths in touch-sensitive toys, with a single terminal for swept-frequency capacitive sensing(Figure~\ref{fig:toys}).

\subsection{Media in Pipes}
After printing, different media can be inserted into pipes to change affordances and capabilities. We consider gas, liquids, solids, particulates, and threadables. 
%
Compressible fluids (``gas'') and incompressible fluids (``liquids'')  inside semi-enclosed pipes can create haptic feedback; pressure can also be sensed to act as an input. Gases can be used as carriers for scents or fog. Conductive liquids like copper paint, when dried, turn pipes into conductive paths and can also fix electronic components in place.
%
%got to here
Filling pipes with solids at print time enables designers to exploit the difference in material properties between pipe and contents. For example, in Printed Optics~\cite{Willis-printedoptics}, different refractive indices between solid cladding material and transparent material inside pipes made light transport along pipes for input and output possible.
%
Particulates, either printed in-place or inserted, can be of varying densities and can contain particles of various and variable size.  A single large particle can be used for display.  Small particles in fluid can provide variable stiffness haptics via jamming \cite{Follmer-jamming}.
%
Threadable media, such as regular wire, electroluminescent wire, or commercial fiberoptic cables, can be threaded through pipes post-printing. Threading such media enables addition of conduction, illumination, or optical sensing. 
% This empowers inexpensive printers: for example, a Printed Optics-style interface can be created on a consumer-grade 3D printer with pipes and fiber optic cable.

>>>>

Open pipes may be used to create electronic circuitry.  For example, an open pipe filled with conductive paint and connected to a battery can provide power to an LED (see Figure \ref{fig:openings}A).
%
Semi-enclosed pipes are capped on one side. Caps can for example be used to create tactile output. By fabricating a cap from a malleable material and attaching a pump to the open ended, haptic feedback can be generated by raising and depressing the flexible cap (see Figure \ref{fig:openings}C).
%
A fully enclosed pipe or cavities are entirely internal to the object.  Fully enclosed pipes may be used as weight modifiers~\cite{Prevost-makeitstand} and for object identification~\cite{Willis-infrastructs}, as in Figure \ref{fig:openings}D.  They can also contain for water or particles that would otherwise fall out.\bjoern{how would you get the particles inside?}

@@

\subsection{Pipe Topologies}
Pipe network topologies enable splitting or mixing of media (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out configurations; we use star topologies to create different conductive paths in touch-sensitive toys, with a single terminal for swept-frequency capacitive sensing (Figure~\ref{fig:toys}).

\subsection{Media in Pipes}
After printing, different media can be inserted into pipes to change affordances and capabilities. We consider gas, liquids, solids, particulates, and threadables. 
%
Compressible fluids (``gas'') and incompressible fluids (``liquids'')  inside semi-enclosed pipes can create haptic feedback; pressure can also be sensed to act as an input~\cite{Slyper-shape}. Gases can be used as carriers for scents or fog. Conductive liquids like copper paint, when dried, turn pipes into conductive paths and can also fix electronic components in place.
%
%got to here
Filling pipes with solids at print time enables designers to exploit the difference in material properties between pipe and contents. For example, in Printed Optics~\cite{Willis-printedoptics}, different refractive indices between solid cladding material and transparent material inside pipes made light transport along pipes for input and output possible.
%
Particulates, either printed in-place or inserted, can be of varying densities and can contain particles of various and variable size.  A single large particle can be used for display.  Small particles in fluid can provide variable stiffness haptics via jamming \cite{Follmer-jamming}. \bjoern{I thought jamming is about particles in a vacuum? why liquid?}
%
Threadable media, such as regular wire, electroluminescent wire, or commercial fiberoptic cables, can be threaded through pipes post-printing. Threading such media enables addition of conduction, illumination, or optical sensing using materials that can easily be procured, but cannot themselves be printed, e.g., on consumer printers with a single material. 
% This empowers inexpensive printers: for example, a Printed Optics-style interface can be created on a consumer-grade 3D printer with pipes and fiber optic cable.

### 2014-04-09 paper/introduction.tex
bjoern's edits through end of tools section

Generating these pipes, however, is not trivial.  For one, most 3D modelling tools do not provide extensive tools for modelling the  {\em interior} of 3D models. Furthermore, careful consideration needs to go into the design and routing of the pipes. Pipes need maximum possible bend radius to ease the insertion of media post-print.  Independent pipes cannot intersect, as this could lead to electrical shorting or other problems.  Pipes should not pass too close to the surface: consumer-grade 3D printers do not guarantee air- or water-tightness of prints, so more than one printed layer may be required to prevent leakage.  Pipes for applications like neon signs must allow a single EL wire to light every edge on the pipe graph (i.e., have an Euler tour).  No tools currently exist for these design tasks. To aid users in working with 3D-printed pipes, we contribute new techniques for interactively routing and generating internal structures, and integrate these tools into a 3D modelling application.

>>>>

Generating these pipes, however, is not trivial.  For one, most 3D modelling tools do not provide extensive tools for modelling the  {\em interior} of 3D models. Furthermore, careful consideration needs to go into the design and routing of the pipes. Pipes need maximum possible bend radius to ease the insertion of media post-print.  Independent pipes cannot intersect, as this could lead to electrical shorting or other problems.  Pipes should not pass too close to the surface: consumer-grade 3D printers do not guarantee air- or water-tightness of prints, so more than one printed layer may be required to prevent leakage.  Pipes for applications like illumunated EL wire signs must allow a single pipe to follow a specific path layout (e.g., spelling out text).  No tools currently exist for these design tasks. To aid users in working with 3D-printed pipes, we contribute new techniques for interactively routing and generating internal structures, and integrate these tools into a 3D modelling application. \bjoern{what's missing here is how our tool achieves the goal. May want to flip this paragraph around - ``we contribute a design tool that addresses various challenges of pipe design makers are likely to encounter. The tool offers 3D point-to-point routing of flexible pipes based on a physical simulation of flexible rods. This process ensures maximum bend angles, which ...'' etc.}

### 2014-04-09 paper/relatedwork.tex
bjoern's edits through end of tools section

Sauron tracks interactions with printed physical mechanisms by placing a camera inside hollow objects, where all components have to be visible to a camera~\cite{Savage-sauron}.  Our technique instead can work with general, solid models and only carves out pipes for interaction where needed. Midas generates fabrication files custom capacitive touch sensors~\cite{Savage-midas}.  This technique, based on circuitboard routing, is limited to planar and unfoldable shapes. Connections are routed on the surface of objects. \systemname can also be used for touch sensing, but routes conductors through the inside of objects.

@@

\subsubsection{Fabrication with Printed Electronics}
Researchers are also investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap}use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals. Our approach does not deposit such materials automatically, but creates structures inside a 3D model that make manual post-print injection of conductive ink possible. 
%Navarette, et al., created a functioning GPS routed in true 3D (not multilayer boards' 2.5D), hand-built using a one-off routing.  \bjoern{need more context for this}
%Our work differs from these because we offer \emph{algorithms and techniques} for \emph{internal}, \emph{true 3D} routing.

@@

One of our path routing algorithms routes a single, continuous pipe through an object based that follows the appearance of an input shape. The technique is inspired by work in generating continuous line illustrations from images~\cite{Wong-continuousline,Bosch-tsp}. Our approach is also based on graph theory (Euler tours), but differs in that it generates 3D geometry instead of 2D illustrations.

@@

Some prior work has also focused on modifying interior parts of 3D objects. 
%Takayama, et al., built a system described in \cite{Takayama-fruit} which allowed users to create internal color maps of the interior of objects, allowing them to see digital ``slices'' of their ``oranges.'' \bjoern{don't understand this} Our work focuses on 3D printable geometry rather than rendering.  
In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}. Infrastructs encodes information using interior cavities that can be sensed using teraherz imaging. Bickel, et al., explored enforcing deformation behavior with multi-material blending inside objects\cite{Bickel-deformation}. Modeling internal chambers is one aspect of our work, however our focus is on interaction rather than balance, identification or deformation.  
%B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}. All these contributions are orthogonal to ours, as our focus is on precisely \emph{removing} material from printed objects, rather than blending materials or modifying components.

>>>>

Sauron tracks interactions with printed physical mechanisms by placing a camera inside hollow objects, where all components have to be visible to a camera~\cite{Savage-sauron}.  Our technique instead can work with general, solid models and only carves out pipes for interaction where needed. Midas generates fabrication files for custom capacitive touch sensors~\cite{Savage-midas}.  Their technique, is limited to planar and unfoldable shapes as it routes connections on the surface of objects. \systemname can also be used for touch sensing, but routes conductors through the inside of objects.

@@

\subsubsection{Fabrication with Printed Electronics}
Researchers are also investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap} use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals. Our approach does not deposit such materials automatically, but creates structures inside a 3D model that make manual post-print injection of conductive ink possible. 
%Navarette, et al., created a functioning GPS routed in true 3D (not multilayer boards' 2.5D), hand-built using a one-off routing.  \bjoern{need more context for this}
%Our work differs from these because we offer \emph{algorithms and techniques} for \emph{internal}, \emph{true 3D} routing.

@@

One of our path routing algorithms routes a single, continuous pipe through an object that follows the appearance of an input shape. The technique is inspired by work in generating continuous line illustrations from images~\cite{Wong-continuousline,Bosch-tsp}. Our approach is also based on graph theory (Euler tours), but differs in that it generates 3D geometry instead of 2D illustrations.

@@

Some prior work has also focused on modifying interior parts of 3D objects. 
%Takayama, et al., built a system described in \cite{Takayama-fruit} which allowed users to create internal color maps of the interior of objects, allowing them to see digital ``slices'' of their ``oranges.'' \bjoern{don't understand this} Our work focuses on 3D printable geometry rather than rendering.  
In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}. Infrastructs encodes information using interior cavities that can be sensed using teraherz imaging~\cite{Willis-infrastructs}. Bickel, et al., explored enforcing deformation behavior with multi-material blending inside objects\cite{Bickel-deformation}. Modeling internal chambers is one aspect of our work, however our focus is on interaction rather than balance, identification or deformation.  
%B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}. All these contributions are orthogonal to ours, as our focus is on precisely \emph{removing} material from printed objects, rather than blending materials or modifying components.
